\huge\bf{Querypoint} : Moving Backwards on Wrong Values in the Buggy Execution

### 2011-04-26 papers/fse2011/reviewResponse.txt

Thank you for reviewing our paper on 'lastChange'. We respond to the
questions raised by the reviewers and then request that the first
reviewer reconsider part of their analysis.

The first reviewer asks:

Q: "The paper presents the work as a general debugging approach. ... how
much of it is made possible through the unique characteristics of
Firefox and Javascript programming? How closely similar could it be
possible to support this idea for a compiled language?"

A: Section 3 of our paper presents a general algorithm and in that
section we give the four prerequisite operations, 1) objectId(), 2)
setPropertyChangeHook(), 3) scopeId(), and 4)
setVariableChangeHook(). A runtime which can support these operations
can run the algorithm; other runtimes cannot. No special feature of
Firefox or of JavaScript is used in this section. Whether a language
is compiled or not does not relate to our technique. We have
implemented this technique for Java as shown in an unreviewed
technical report: http://infoscience.epfl.ch/record/164417. This
technique should apply to C, C++, C#, Pascal, PHP, Ruby, and so on.

The second reviewer asked many important questions about our small user study:

Q: "Who were the developers?"

A: Unfortunately it would not be practical to give the names of our study
particiapants. We assume the reviewer is asking for more information
about the skill of the developers, which we addresss in the remaining
answers.

Q: "How much experience did they have?"

A: These were all skilled developers with several years of experience in
debugging complex programs.

Q: "Are they researchers?"

A: Yes of course. Practically speaking, we cannot hire 'average'
developers to engage in a lengthy study of such a complex subject as
debugging code. A small study with "folks down the hall" is what is
practical.

Q: "Do they know the concept of a data dependency?"

A: Yes. However, our prototype does not require users to know this concept.

Q: "Do they work on programming tools?"

A: No.

Q: "How complex were the test applications?"

A: These were simple applications and they are available on the Web site
we listed in footnote 4,
http://ltiwww.epfl.ch/~mirghase/lastchange-userstudy.

Q: "We're they already familiar with Firebug and its debugging tools?"

A: Of course, most JavaScript developers use Firebug.

Q: "How was the task explained to them?"

A: The instructions were provided on the Web pages listed in footnote 4,
http://ltiwww.epfl.ch/~mirghase/lastchange-userstudy/group1.html.

Q: "We're they given time limits?"

A: As described in the Web pages listed in footnote 4, the introduction
and practice sessions where not timed, then two time trials were
performed. That is how we arrived that the numbers in Figure 12.

Q: "How were they recruited?"

A: We asked for volunteers from among people in our building who we
knew wrote JavaScript.

Q: "What counted as a "step" in Figure 12)."

A: Oops we should have explained that a step in Figure 12 means a button
push, either single stepping the debugger or running a querypoint.

Q: "The sample size of 4 was extremely small, preventing the authors from
really avoiding learning effects (they did counterbalance exposure to
each program)."

A: We did counterbalance amoung users but as you say the small sample
makes this ineffective. We would like to point out that learning
effects stongly bias against our technique. As the reviewers surely
experience, debugging takes a long time to master. New debugging
techniques take a long time to learn. That our test subjects were able
to succeed in less than an hour is itself the most important result of
our small study.

Finally we ask reviewer #1 to reconsider section 4 of our paper and
the related work section discussing reverse watchpoints.

We are very surprised at the claim of reviewer #1 that we don't build
on top of existing debug technology. Section 4 of our paper details
how we built lastChange on top of the Firebug debugger. Perhaps we
should have been clearer: by "breakpoint technology" we mean all the
different ways debuggers pause execution when a special event
occurs. It includes line step, function enter/exit, variable/property
access/change (watchpoint), error/exception, conditional breakpionts,
and changes on graphical changes. (see for example our ref. 1). Our
ability to build upon existing debugger technology is essential to the
practicality of our approach.

We do not believe the claim of the first reviewer that a developer can
"manually perform a lastChange query in relatively few steps." In fact
our user study, as small and limited as it may be, provides direct
evidence against this claim. Moreover our approach can be applied
recursively, to obtain the lastChange to a value found on a
lastChange.

Reviewer #1 points to a combination of watchpoints and the "drop in
frame" feature in Eclipse. Watchpoints trigger on data value changes;
"Drop in frame" allows for shorter re-execution cycles like the local
replay we explained in section 4.5. The fundamental problem with these
techniques is that they only work in the 'forward' direction. If we
land on a watchpoint after a 'drop in frame' operation, then is this
the last change or just one of a hundred changes? This fundamental
problem is why papers cited as our references 4, 5, 6, 8, 11, 12, 13,
and 15 exist: developers want to go backwards. That is why we called
our paper "lastChange". As cited in our related work section, our
approach is superior to even reverse watchpoints, a technology created
to go beyond normal watchpoints. We could of course include a
discussion of normal watchpoints.

Again we thank the reviewers. We believe small changes to the text
based on this feedback would clarify and improve the paper.

### 2011-06-07 trunk/papers/fse-tool2011/sig-alternate.tex

\title{Debugging by {\huge\bf\textit{lastChange}}}

@@

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Salman Mirghasemi\\
       \affaddr{\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland}\\
       \email{salman.mirghasemi@epfl.ch}
% 2nd. author
\alignauthor
John J. Barton\\
       \affaddr{IBM Research - Almadan}\\
       \email{bartonjj@us.ibm.com}
% 3rd. author
\alignauthor 
Claude Petitpierre\\
       \affaddr{\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland}\\
       \email{claude.petitpierre@epfl.ch}
}

@@

\begin{abstract}
%Developers often seek the origins of wrong values they see in their
%debugger. Their search must be backwards in time: the code causing the
%wrong value executed before the wrong value appeared. Searching with
%breakpoint- or log- based debuggers demands persistence and
%significant experience with the application being debugged. 
We introduce a new, practical feature for debuggers called \textit{lastChange}, 
which automatically locates the last point that a variable or an object property 
has been changed. Starting from a 
program halted on a breakpoint, the \textit{lastChange} solution
applies queries to the live program during re-execution, recording the call stack 
and limited program state each time the property value changes.
When the program halts again on the breakpoint, the recorded information 
can be shown to the developer.
As a proof of this concept, we developed \textit{Querypoint}, a prototype
which enhances the popular Firebug JavaScript debugger with
 the \textit{lastChange} feature and studied users applying the prototype to 
some test cases.
 The approach used in implementing 
\textit{lastChange} combines the flexibility of breakpoint debugging
with the expressive power of log-based query debugging.  Contrary to
other replay-based approaches, which require exactly the same
re-executions (deterministic executions), our new approach only requires \textit{bug 
reproducibility}, meaning a test case is available which reproduces the bug and a 
way to halt execution reliably after the reproduction.

\end{abstract}

@@

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

@@

In log-based debugging, developers select data to be collected by
inserting statements for all points of possible change.  While in
breakpoint-based debugging, the whole program state is available to
developer, in log-based debugging, developer has to decide what data
should be collected when inserts the log statement. It is very common
that the developer has to repeat this step several times due to
insufficient collected data, or to wait a long time because too much
data is recorded. Once adequate data is collected, it still
requires analyzing and understanding. Developers usually end up in
dealing with long log files and analyzing huge amounts of collected data. 
Neither approach efficiently assists the developer in finding origins to a wrong
value.

@@

The rest of the paper is organized as follows. First, we demonstrate
the \textit{lastChange} usage on a simple example with the comparison
to breakpoint debugging. Section 3 presents \textit{lastChange}
algorithm. In section 4, we explain the details of the JavaScript
prototype implementation. We discuss the effect of non-determinism on
the \textit{lastChange} results in section 5. The user study results 
are presented in section 6. %TODO update it according to new Section 6 contains an evaluation of\text{lastChange} functionality.

@@

\begin{figure}[htp]
\centering

\includegraphics [width=.48\textwidth]{5-example-points.pdf} 
\caption{The examined points before locating the defect. The arrow represents the logical forward progress of the program. Three actual executions are superimposed on this arrow. All three stop at the reproduction point indicated by circle 1. After the first execution, the developer asks for lastChange as described in section \ref{sec:introExample}, yielding information  indicated by circle 2. After the second execution, another lastChange query causes a third execution, yielding information indicated by circle 3.}
\label{fig:example-points}
\end{figure}

@@

\section{lastChange Algorithm}

The \textit{lastChange} algorithm is based on program re-execution of
a program halted on a breakpoint. The algorithm starts when developer
examines the program state at a breakpoint hit and asks for the
\textit{lastChange} of a value. The breakpoint hit becomes the
\textit{reproduction point}. Debugger sets hooks (a callback
function dependent upon the underlying runtime) on all instructions
that might be the result of \textit{lastChange} query. Then the
debugger re-executes the program and every time a hook hits it
checks for a \textit{change event}. In the case of a change, it stores
part of the program state values.  Once the execution reaches the
reproduction point, it analyzes the collected data and shows the
result.  The program state at the execution point of the last change
event is the \textit{lastChange}.

As we described in the preceding section, a \textit{lastChange} query can
be performed on the result of another \textit{lastChange} query. If we
name the reproduction point \textit{R}, we can write the first
\textit{lastChange} in the introductory example in this form:
\textit{lastChange(R, myObject.myProperty)}. It means that this query
is defined at \textit{R}. If we name the result of this query
\textit{L}, we can write the second \textit{lastChange} in this form:
\textit{lastChange(L, myCondition.value)}. In this way, a sequence of
\textit{lastChange} queries with any length can be defined.

\textit{lastChange} can be called on object property, on a variable value, 
or on the results of a \textit{lastChange}. Moroever, common data structures such as
arrays and hashmaps are also supported as special cases of \textit{lastChange} on object property.
We explain each case in the following subsections.

\subsection{{\large\bf\textit{lastChange}} on Object Property}
To simplify the algorithm explanation and defer technical details, we
define two basic operations and later we explain the details of these
two operations. The first operation is \texttt{objectId()}: given a
JavaScript object it returns an integer as its identifier. This
identifier is unique to the object during one execution.  By using an
object id instead of an object reference we allow the garbage
collector to reclaim the space for dead objects just as it would in the
absence of the debugger. The second
operation is \texttt{setPropertyChangeHook()}: given a function and a
string, the function is called whenever a property changes and its
name matches the string. For example, if the string is 
\texttt{foo}, changes to \texttt{bar.foo} or \texttt{baz.foo} would
call the function.  The callback function receives a reference to the
owner of \texttt{foo}.

To see how these functions work, suppose the developer asks for the
last change of \texttt{bar.foo} at the reproduction point in a
program. The debugger calls \texttt{setPropertyChangeHook()} with
\texttt{foo} as the property name and re-executes the
program. Whenever \texttt{foo} changes and the callback function is to
be called, debugger first calls \texttt{objectId()} on the
\texttt{foo} owner object. Then it stores this owner id, the stack
frame locations, and other state values in scope at the call point. 
Then the callback returns to continue the execution. Thus the query is not 
a breakpoint in the sense of pausing for user interaction, but breakpoint 
technology can be used to implement the query.
Whenever the execution reaches the reproduction point the debugger
looks at the history of \texttt{foo} changes and finds the last
\texttt{foo} change with the same object id as \texttt{bar} id at the
reproduction point. Figure~\ref{fig:foo-changes1} shows the list of
property \texttt{foo} change events in a hypothetical
execution. \texttt{bar} id at the reproduction point is 1010, so the
last change of \texttt{bar.foo} is the fourth column.

\begin{figure}[htp]
\centering 
\includegraphics [width=.48\textwidth] {6-foo-changes1.pdf} %[width=.48\textwidth][scale=0.75]
\caption{A hypothetical list of  change events for a property \texttt{foo}. 
Each change event adds a column with the id of the object changed, the call stack,
and some program state such as local variable values. 
At the reproduction point  we determine which id  corresponds to 
object \texttt{bar} and read out column 4,  the last
  change of \texttt{bar.foo}. Column 3 is also a change of
  the object we want to study, id 1010, but it is not the last change;
  Column 5 is also a change of a property \texttt{foo} but it is not for
  the object we are interested in.}
\label{fig:foo-changes1}
\end{figure}

\subsection{{\large\bf\textit{lastChange}} on Variable} 
In JavaScript, every frame has a scope chain and every available
variable in the frame comes from one of the scopes in the frame's
scope chain. Once the developer asks for the last change of a variable with name 
\texttt{foo} at the reproduction point, the debugger first determines the
variable's scope as follows: it iterates over the scopes in the scope
chain and the first scope which has a variable with the same name is
the variable's scope. There are five different scope types: global, local,
closure, \texttt{with} and \texttt{catch}. We explain these
cases in two groups.

\subsubsection{global and \texttt{with} scopes}
Global scope is the most outer scope in the scope chain and it is
also referred to as the global object (the \texttt{window}
object in Web pages). This scope is a regular JavaScript object and therefore every
global variable is a property of global object. Similarly,
 \texttt{with} scopes are also regular JavaScript objects. A \texttt{with}
scope is created by a \texttt{with()} block with an object as the
parameter. Every property of this parameter object is available inside the
block as a variable. \textit{lastChange} treats the case where variable's scope
is global or \texttt{with}, like it does on an object
property.

\subsubsection{local, closure and \texttt{catch} scopes}
Local scope refers to the most nested scope in the scope
chain which contains the local variables. Closure scope refers to the 
scope which is created for a nested function and contains variables 
defined in the outer block. Catch scope is the scope created in the catch
block of try-catch statements and contains the exception
variable. These scopes are not necessarily regular JavaScript
objects.  
Therefore, to track changes to a variable in these scopes we
employ a different approach.

\begin{figure}[htp]
%\begin{verbatim}
\includegraphics [width=.48\textwidth] {9-scopes.pdf} %[width=.48\textwidth][scale=0.75]

%\lstset{basicstyle=\scriptsize}
%\begin{lstlisting}[language=myLang, framerule=0pt]
%  function f(){
%    var x = 0;	
%    x++;
%    ...
%    if (!stop) f();
%  }

%Sample Trace:

%   f()
%A  |  x changes, scope 1 
%   |  f()
%   |  |  x changes, scope 2
%   |  |  f()
%B  |  |  |  x changes, scope 3 
%C  |  ... , scope 1
%\end{lstlisting}
%\end{verbatim}
\caption{A recursive call trace illustrating the scope
  id. The lines in the bottom half of the diagram simulate a trace of the change events for the variable \texttt{x} as the function \texttt{f()} calls itself. Each call creates a scope; eventually when the variable \texttt{stop} is changed by an external process we return from the recursion. The lines marked A, B, and C are discussed in the text.}
\label{fig:recursive}
\end{figure}

Having the scope chain and the source code, we can map every scope to % perhaps needs more explanation, how?
a code block, enclosed the executing code. In
JavaScript, a code block can be identified by the file url and  
the block's first instruction program counter. Given this information, the debugger is
able to recognize the code block in loaded scripts or once it is loaded.
Similar to \textit{lastChange} on object property, we define two basic operations: 
\texttt{scopeId()} -- which returns an integer given a scope -- and \texttt{setVariableChangeHook()} -- which calls its first parameter, a callback function, when a 
variable in its second parameter, a code block, matches the name given in the third parameter.

Figure~\ref{fig:recursive} illustrates how the \texttt{scopeId()} operation separates instances 
of a variable in different scopes having the same name.
If we ask for the last change on \texttt{x} at point labeled \texttt{C} in scope with id 1, we want the change at line \texttt{A} in scope 1, not 
the change at the line marked \texttt{B}, where a variable named \texttt{x} in scope 3 is changed.

If the developer asks for the last change of variable \texttt{foo} at the 
reproduction point in a program, debugger calls 
\texttt{setVari- ableChangeHook()} with the variable's defining
block and name as parameters and re-executes the
program. Whenever \texttt{foo} changes and the callback function is to
be called, debugger first calls \texttt{scopeId()} on the
variable's scope. Then it stores this scope id, the stack
frame locations, and other state values in scope at the call point.
Whenever the execution reaches the reproduction point the debugger
looks at the history of \texttt{foo} changes and finds the last
\texttt{foo} change with the same scope id as the variable's scope id
at the reproduction point (Figure~\ref{fig:foo-changes2}).

\begin{figure}[htp]
\centering 
\includegraphics[width=.40\textwidth]{7-foo-changes2.pdf}
\caption{A hypothetical list of variable \texttt{foo} change events. Each column of the list indicates a change event; for each
change event the scope id returns by \texttt{scopeId()} is recorded along with call stack and program state information. 
   The last column having the scope id \texttt{foo} at the reproduction point indicates 
  the last change.}
\label{fig:foo-changes2}
\end{figure}

\subsection{{\large\bf\textit{lastChange}} on {\large\bf\textit{lastChange}}}
The \textit{lastChange} algorithm records changes by id (either object or scope id), then reads out the last change when we
arrive at the reproduction point and discover the id of the requested value.
When we perform \textit{lastChange} based on a previous \textit{lastChange}, the query algorithm 
must retain additional information. Consider the following example:

\lstdefinelanguage{myLang2}
{morekeywords={point, lastChange}}

\lstset{basicstyle=\small}
\begin{lstlisting}[frame=single, language=myLang2] %, framerule=0pt]

 point A : the reproduction point 
 point B : lastChange(A, bar.x) 
 point C : lastChange(B, baz.y) 
\end{lstlisting}

%\begin{center}
%\textit{
% }
% \end{center}

where point A is a breakpoint, point B is the last change of the object property \texttt{bar.x} at point A, and
point C is the last change of the object property \texttt{baz.y} at point B. 
The object referenced by \texttt{baz} changes upon re-execution. Therefore when the developer
asks for the last change of \texttt{baz.y}, we need to track objects named \texttt{baz}
at changes of \texttt{bar.x} and changes to objects named \texttt{y}. Then, at the reproduction point,
we need to work out which \texttt{baz} the developer wanted, then select the last change of that \texttt{baz.y}.
Figure~\ref{fig:lastchange-lastchange} illustrates the extra row of data (tracking of \texttt{baz} objects 
at \texttt{x} change events) 
and how the id values allow the last change of \texttt{baz.y} to be worked out.

In the general case we perform dependency analysis as outlined in 
Figure~\ref{fig:dependency-analysis} to
create the list of additional data (object id or scope id) to be collected at a change
event. The process can be repeated to cascade \textit{lastChange} arbitrarily deep.

\begin{figure}[htp]
\centering 
\includegraphics[width=.48\textwidth]{8-lastchange-lastchange.pdf}
\caption{The list of change events stored for locating point B, the
  \textit{lastChange} of \texttt{bar.x} at the reproduction point, and
  point C, the \textit{lastChange} of \texttt{baz.y} at point B.}
\label{fig:lastchange-lastchange}
\end{figure}

\begin{figure}[htp]
%\centering 
%\begin{algorithm}
\line(1,0){240}
\begin{algorithmic}%[frame=single]
\FOR{$q$ in $lastChange$ queries} 
 \FOR {$p$ is defined at the $q$ result}
   \IF {$p$ is a $lastChange$ on object property} 
     \STATE the property {\bf owner id} must be stored at $q$ change events. 
   \ELSIF {$p$ is a $lastChange$ on variable} 
     \STATE the variable {\bf scope id} must be stored at $q$ change events.
	 \ENDIF 
 \ENDFOR 
\ENDFOR

\end{algorithmic}
\line(1,0){240}
%\end{algorithm}
\caption{\textit{lastChange} queries dependency analysis.}
\label{fig:dependency-analysis}
\end{figure}
%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented\footnote[1]{http://code.google.com/p/querypoint-debugging} it in an extension to the Firebug
JavaScript debugger\footnote[2]{http://getfirebug.com}. %\cite{Firebug}. %\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser. %\cite{Firefox}. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface \cite{JSD} and 
\textit{Querypoint} is developed over this interface. Our prototype implements the four primitive operations in JavaScript 
using techniques which are cumbersome and comparatively slow to execute. However the JavaScript prototype is
easy to explore, change, and share with others for feedback. A professionally useful debugger 
would implement these primitive operations within the JavaScript engine.

\subsection{{\normalsize\bf\texttt{objectId()}} Operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
%\footnote[3]{The property name we used in the prototype is longer and more specific to reduce the chance of name conflicts.} .
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. 
%\footnote[3]{This use of a regular object property  might change
%the program behaviour. For example it will appear as an extra property in 
%\texttt{for(property in object)} loops. 
%The next version of Firefox (https://developer.mozilla.org/en/JavaScript/Reference\\/Global\_Objects/Object/defineProperty) provides \texttt{defineProperty}, 
%allowing a property to be added which is invisible to enumeration.} Note that \texttt{\_objectId} is
%not set for all objects but only those objects that need an id.

\subsection{{\normalsize\bf\texttt{setPropertyChangeHook()}} Operation}
The Firefox JavaScript engine supports watching property changes in an
object. Every object has a function \texttt{watch(propName, callback)}
which receives two parameters, a property name and a function.  Whenever
the property with the given name changes, the \texttt{callback} function is
called. The hook set by this function remains enabled even if the
property is deleted and defined again.

For our purposes, the \texttt{watch()} function only covers the case
of global object properties. At the beginning of execution, no object
excepting the global object and its predefined properties is available.  For 
our \textit{lastChange} prototype, we created a version of
\texttt{setPropertyChangeHook()}.  The basic strategy is to get a reference to the object just
after its creation, then use \texttt{watch()} function to monitor
property changes in the object. Setting a flag\footnote[3]{DISABLE\_OBJECT\_TRACE defined in jsdIDebuggerService.idl} into the Firefox 
JavaScript engine, we can get the file URL and the line number for each object
creation (e.g., myFile.js, line 24). We set a breakpoint on this line
and parse the source code to determine which object was created.

The only data we have is the object creation location including the
file url and the line number and the goal is to get a reference to
this object. Although in most regular cases we have only one
statement and one object creation, there are cases where more
objects are created in the line. There is no simple way to
recognize the interesting object among these new objects. So instead
of one object, we monitor all new objects created in the line.

An object might be created by one of these statements: object
literal(\texttt{\{...\}}), \textit{new} operator (\texttt{new 
constructor()}) or function definition statement (\texttt{function()}). 
By parsing the source code we can recognize the statements that
create an object. The next step is getting a reference to the new object.

The new object can be assigned to an object property or a variable by
an assignment (\texttt{=}). In these cases we keep the assignee statement 
at the left side. The idea is that we create a list of assignee statements 
that the new objects are assigned to. We set a hook on the creation
line. Once the hook hits, we evaluate the assignee statements. Then 
we do stepping(step-over) and after each step we evaluate the
statements. Every statement which has a new value, we consider the new
value (if it is an object) as a new object. For example in Figure
~\ref{fig:objectCreation}(a), if the creation line number is 20, we
have only one statement which creates a new object and it is assigned
to \texttt{x.y}.

The new object can also be set as the property of a parent object 
(or array) inside an object (or array) literal. This case is also treated 
similar to previous case. The only difference is that the full path of property 
from the root parent in the local scope must be considered as the assignee statement. For
example in Figure~\ref{fig:objectCreation}(b), if the creation line
is 20, we keep \texttt{parent.child}. In this case stepping
must be continued until the end of the object literal (line 22) for getting a reference
to the created new object.

In cases where the new object is passed as an 
argument to a function (Figure~\ref{fig:objectCreation}(c)), we use step-in instead of step-over. 
To get a reference to the new object
it is enough to evaluate the corresponding argument inside the function. The other cases
where the program does not keep a reference to the created object (e.g., when a function object is
just called after its creation), are not in our interest.

%(or \texttt{this} if the object is being constructed by \texttt{new} statement)

Although this approach is successful in many ordinary cases, we can imagine 
cases where a more comprehensive analysis needed for the correct behaviour. 
Consider the case where the new object is assigned to an expression like \texttt{a[++i]}. 
% (where i is a variable). 
%This expression is kept in the list                      %(Figure~\ref{fig:objectCreation}(d))
%of assignee statements to be evaluated later for getting a reference to the new object. 
Obviously, evaluating this statement doesn't return
a reference to the new object. Our prototype implementation does not handle these kinds of unusual cases yet.

Throughout this section we implicitly assumed that the object will be created at 
the same location in the next execution. If the assumption is
not true, once the execution reaches the reproduction point, it
reveals that the object has been created in a different location. This time,
prototype re-executes the program considering both locations as possible 
object creation locations.

\begin{figure}[htp]
%\begin{verbatim}
\lstset{basicstyle=\scriptsize}
\begin{lstlisting}[frame=single, language=myLang]% framerule=0pt]

(a) Case 1:
 20  x.y = new MyClass();

(b) Case 2:
 19  parent = {
 20    child : {
 21      x : 5
 22  }}

(c) Case 3:
 20  myFunction({myProperty:5});

\end{lstlisting}

%(d) Case 4:
% 20 a[++i] = new myConstructor();

%\end{verbatim}
\caption{Examples of different cases in getting references to created JavaScript objects.}
\label{fig:objectCreation}
\end{figure}

\subsection{{\normalsize\bf\texttt{scopeId()}} Operation}
The prototype sets a breakpoint at the beginning of all code blocks needing
a scope id. The scope id is kept as a variable with name 
\texttt{\_scopeId} in the scope. Whenever the hook is hit,  meaning a new scope
is created, \texttt{\_scopeId} is set by calling JavaScript's dynamic compilation 
function \texttt{eval()}. For example, executing \texttt{eval("var \_scopeId = 10")} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope. \texttt{scopeId()}
operation returns the value of \texttt{\_scopeId} in the scope.

\subsection{{\normalsize\bf\texttt{setVariableChangeHook()}} Operation}
Variables defined in local, closure, and catch scopes are only changed in
their scope; that includes the defining scope and scopes nested whithin them.
For example in Figure~\ref{fig:js-closure}, variable \texttt{foo} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line where the variable is defined. In Figure~\ref{fig:js-closure}, 
if the execution is paused at line 17 and the
last change of \texttt{foo} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different: \texttt{foo} in 
\texttt{childOne} is a local variable
but in \texttt{childTwo} it is a closure variable.

\begin{figure}[htp]
%\begin{verbatim}
\lstset{basicstyle=\scriptsize}
\begin{lstlisting}[frame=single, language=myLang]%, framerule=0pt]

10  function main(){
11    var foo;
12    foo = ...;
13    function parent(){
14      foo = ...;
15      function childOne(){
16        var foo;
17        foo = ...;
18      }  
19      function childTwo(){
20        foo = ...;			      
21      }
22    }  
23  }    
\end{lstlisting}
%\end{verbatim}
\caption{Sample JavaScript code demonstrating local and closure variables.}
\label{fig:js-closure}
\end{figure}

\subsection{Re-Execution, Reproduction Point and Data Collection}
\textit{Querypoint} needs a test case to reproduce the
execution and conditions to correctly recognize the reproduction point. 
Although both elements can be directly provided by developer, \textit{Querypoint}
is also able to automatically create them from the first execution.

To replay execution, \textit{Querypoint} keeps track of breakpoint hits and single steps. For example, 
if the developer queries \textit{lastChange} at the third hit of breakpoint \textit{b}, in
re-execution, the third hit is recognized as the reproduction point.

The \textit{Querypoint} prototype supports two mechanism
for automatic re-execution: callstack-reproduction and record-replay. In callstack reproduction the function from
the earliest frame of the call stack is called with the same parameters. The idea
behind this mechanism is that many bugs in web pages can be reproduced by re-firing
an event like clicking on a button. The record-replay execution uses two phases. In the record phase, it 
stores the initial page url and the events and parameters corresponding to user actions. In the replay phase, it opens
the same url and simulates events as if they were user actions. The callstack reproduction mechanism provides
shorter re-execution cycles while the record-replay is more accurate about the
initial state.

In addition to the data collected at every change event for identifying the \textit{lastChange}
result, \textit{Querypoint} partially stores values in program state. There is a trade-off between the amount of data collected at every change event and the number of re-executions. If developer asks for 
some values which have not been stored, \textit{Querypoint} re-executes and collects the requested data.

% Discuss what happens when assumptions changes, like object creation location.

\section{Reproducible Non-Deterministic Execution}
We claim that the only prerequisite for \textit{lastChange}  is bug-reproducibility. 
A bug is \textit{reproducible} for a developer when the developer can
start from a given initial state, operate on the program with a
list of actions, and reproduce the symptoms of the bug. The details of
the execution can change each time we re-execute the buggy program,
but the buggy result is the same.  All modern debuggers in wide use that we know 
about rely on reproducible but non-deterministic execution for simple practical reasons: 
developers must reproduce a bug to study it and modern execution environments 
are not deterministic.  
% In this section we discuss non-determinism and \textit{lastChange}. However we want 
%to emphasize that when developers have to think about non-determistic execution 
%it usually means that bugs are not reproducible. 
%The reproducibility of the bug
%means that the defect is very unlikely to depend on the order of
%events during the execution. Conversely, if the defect depends on the order of execution, 
%it will likely fail to reproduce on some executions. 
By relying on reproducibility but 
not requiring deterministic execution, \textit{lastChange} works on the same range of cases.

\subsection{{\large\bf\textit{lastChange}} Result Consistency}
\label{sec:resultConsistency}

Each time we re-execute a non-deterministic program, 
the details of execution instruction order may change. 
For example, if we record the source code lines every time
a conventional watchpoint hits, the record may differ each time we
re-execute. But \textit{lastChange} does not compare values across 
executions. Rather it analyzes all of the
change events in a single execution. 
Neither the data gathering nor the analysis require deterministic execution.

>>>>

%\title{Debugging by {\huge\bf\textit{lastChange}}}
\title{\huge\bf{Querypoint} : Moving Backwards on Wrong Values in the Buggy Execution}

@@

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Salman Mirghasemi\\
       \affaddr{\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland}\\
       \email{salman.mirghasemi@epfl.ch}
% 2nd. author
\alignauthor
John J. Barton\\
       \affaddr{IBM Research - Almaden}\\
       \email{bartonjj@us.ibm.com}
% 3rd. author
\alignauthor 
Claude Petitpierre\\
       \affaddr{\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland}\\
       \email{claude.petitpierre@epfl.ch}
}

@@

\begin{abstract}

\end{abstract}
Developers often seek the origins of wrong values they see in their
debugger. Their search must be backwards in time: the code causing the
wrong value executed before the wrong value appeared. Searching with
breakpoint- or log- based debuggers demands persistence and significant experience with the application being debugged.

\textit{Querypoint}, is a Firefox plugin which enhances the popular Firebug JavaScript debugger with
 a new, practical feature called \textit{lastChange}, which automatically locates the last point that a variable 
or an object property has been changed. Starting from a program halted on a breakpoint, the \textit{lastChange} solution
applies queries to the live program during re-execution, recording the call stack and limited program state each time 
the property value changes. When the program halts again on the breakpoint, the recorded information can be shown to the developer.

@@

@@

In log-based debugging, developers select data to be collected by
inserting statements for all points of possible change.  While in
breakpoint-based debugging, the whole program state is available to
developer, in log-based debugging, developer has to decide what data
should be collected when inserts the log statement. It is very common
that the developer has to repeat this step several times due to
insufficient collected data, or to wait a long time because too much
data is recorded. Once adequate data is collected, it still
requires analyzing and understanding. Developers usually end up in
dealing with long log files and analyzing huge amounts of collected data. 
Neither approach effectively assists the developer in finding origins to a wrong
value.

@@

\section{Related Work}
\label{sec:relatedWork}
%whyline, firecrystal

Tools developed to enhance developers' movement on the buggy execution can be classifed in two groups: replay-based and logging-based.

%The algorithm we presented in this paper obtain information about
%the execution state logically earlier in the control flow by querying 
%the executioin state during replay using the technology of 
%conventional breakpoint debuggers.  
The \textit{lastChange} approach resembles the operational model of replay-based debugging 
and the query approach of logging-based
debugging.  Replay-based approaches capture limited data during
execution and replay the buggy execution to reach past points. In
contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution, then query the data to 
inform the developer. Replay-based
approaches impose much less runtime overhead (about two orders of
magnitudes) comparing to logging-based approaches. However, developer
has to re-execute the buggy execution several
times. \textit{lastChange} collects data on re-execution by queries
selected by developer interaction with the debugger. Therefore it has 
the selectivity of the replay-based approaches that improves performance, 
% but we have 
and the flexibilty of the queries so it does not require deterministic replay.

Logging-based debuggers suffer from different issues. First, the recording
step is time expensive and it should be repeated in case of changes in
program. Second, the execution log cannot fully replace the live
execution. There are other aspects of execution (e.g., program user
interface, file system, database tables, etc.) which are also
important in debugging and are not available to the developer in 
logging-based debuggers. Third, querying collected data (e.g., to restore
the program state at a certain point) may not be efficient enough for
debugging realistic programs. Comparing to logging-based debuggers our approach 
has little upfront cost and more flexibility. The developer can start debugging just 
after reproducing the bug without a capturing step.  Changing inputs or
environment settings and re-executing to investigate the bug works as
in conventional breakpoint debuggers.

@@

@@

\subsection{Re-Execution, Reproduction Point and Data Collection}

\section{Limitations and Future Work}

\section{Reproducible Non-Deterministic Execution}

\subsection{Combination of {\large\bf\textit{lastChange}} and Breakpoint Debugging}

\section{Conclusion}

\section{Acknowledgments}
